\documentclass[10pt]{article}

\usepackage[margin=0.82in]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

\title{QuorumScope: A Bounded Consistency Testing Workbench for Replicated Registers}
\author{Tazeem Mahashin\thanks{Rensselaer Polytechnic Institute}}
\date{\today}

\definecolor{qblue}{HTML}{1F4E79}
\definecolor{qgreen}{HTML}{0F766E}
\definecolor{qred}{HTML}{A83224}
\definecolor{qgray}{HTML}{F4F6F8}
\definecolor{qline}{HTML}{CCD5DA}

\newcommand{\qs}{QuorumScope}
\newcommand{\firstack}{first-ack}
\newcommand{\quorum}{quorum}
\newcommand{\cmd}[1]{\texttt{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\captionsetup{font=small,labelfont=bf}
\setlist[itemize]{leftmargin=1.35em,itemsep=0.2em,topsep=0.25em}
\setlist[enumerate]{leftmargin=1.45em,itemsep=0.2em,topsep=0.25em}
\setlength{\parskip}{0.35em}
\setlength{\parindent}{0pt}
\emergencystretch=2em

\begin{document}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\maketitle
\setcounter{footnote}{0}
\renewcommand{\thefootnote}{\arabic{footnote}}

\begin{abstract}
\qs{} is an artifact for testing consistency behavior in a replicated single-key register under
network partitions. It compares an intentionally unsafe \firstack{} protocol with a majority
\quorum{} protocol by running both protocols over the same modeled schedules, extracting
operation histories, checking linearizability, shrinking counterexamples, and preserving
regression fixtures. The report focuses on the research artifact: the model, the evidence
pipeline, the checker, the generated and exhaustive search processes, and the findings produced
by the current implementation. In the current snapshot, \firstack{} produces stale-read
linearizability violations across curated, generated, benchmark, corpus, and tiny exhaustive
surfaces. \quorum{} produces zero linearizability violations in those bounded corpora under the
modeled assumptions, while reporting unavailable operations as an explicit availability cost.
This is bounded evidence for the implemented model, not a proof of quorum correctness in
arbitrary distributed systems.
\end{abstract}

\section{Research Question and Contributions}

The artifact asks a narrow systems question: when a replicated register is partitioned, can a
protocol that returns after the first reachable replica response produce a successful stale read
that violates linearizability, and can the artifact discover, minimize, replay, and explain that
failure without relying on a hand-curated example?

\qs{} contributes four pieces of evidence machinery:

\begin{enumerate}
  \item A deterministic partition simulator that turns bounded schedules into operation histories
  for competing protocols.
  \item A single-register linearizability checker that returns legal-order witnesses for safe
  histories and stale-read witnesses for unsafe histories.
  \item A search and shrinking path that finds generated counterexamples and reduces them to
  small replayable scenarios.
  \item A tiny exhaustive explorer and regression corpus that make bounded coverage and fixture
  drift visible instead of implicit.
\end{enumerate}

\begin{center}
\fcolorbox{qline}{qgray}{\begin{minipage}{0.92\linewidth}
\textbf{Claims and non-claims.}
\begin{itemize}
  \item \textbf{Claim:} \firstack{} exhibits stale-read linearizability violations in the
  reported curated, generated, benchmark, corpus, and tiny exhaustive evidence surfaces.
  \item \textbf{Claim:} \quorum{} produced zero linearizability violations in the declared
  bounded corpora under the current single-register partition model.
  \item \textbf{Claim:} \quorum{} pays an availability cost; unavailable operations are counted
  and reported.
  \item \textbf{Non-claim:} this is not a general proof of \quorum{} correctness.
  \item \textbf{Non-claim:} this is not a production database, Raft/Paxos implementation, real
  network simulator, durability model, or Byzantine-fault model.
\end{itemize}
\end{minipage}}
\end{center}

\section{System Model}

The model is a deterministic replicated register with one key, a finite set of replicas, and
client read/write operations. A scenario is a bounded sequence of operation steps and topology
steps. Topology steps change replica reachability by creating healed or partitioned groups.
Operation steps execute under a selected protocol and record invocation time, response time,
status, consulted replicas, and observed or written value.

Unavailable operations are first-class evidence. They are not passed to the linearizability
checker as successful completed operations, but they are reported because they describe the
safety/availability tradeoff. A protocol that avoids stale reads only by refusing minority-side
operations should make that refusal visible.

The artifact currently compares two register protocols:

\begin{itemize}
  \item \textbf{\firstack{}.} A read or write completes after the first reachable replica response.
  This deliberately weak rule can return stale data from a minority partition.
  \item \textbf{\quorum{}.} A read or write requires a majority of replicas. In the modeled
  stale-read schedules, this prevents minority reads from completing successfully.
\end{itemize}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\small,
  box/.style={draw=qline, rounded corners=3pt, fill=qgray, align=center, minimum width=4.2cm, minimum height=1.15cm},
  redbox/.style={box, draw=qred!60, fill=qred!6},
  greenbox/.style={box, draw=qgreen!65, fill=qgreen!7},
  arrow/.style={-{Latex[length=2.3mm]}, thick, draw=qblue}
]
  \node[greenbox] (maj) at (0,0) {\textbf{Majority partition}\\\texttt{n3, n4, n5}\\write \texttt{op2}: commit \texttt{v1}};
  \node[redbox] (min) at (7.1,0) {\textbf{Minority partition}\\\texttt{n1, n2}\\read \texttt{op3}: returns \texttt{v0}};
  \node[box, minimum width=8.6cm] (check) at (3.55,-2.05) {\textbf{Linearizability checker}\\real-time order requires \texttt{op2} before \texttt{op3};\\single-register oracle expects \texttt{v1}, observes \texttt{v0}};
  \draw[dashed, thick, draw=qline] (3.55,0.95) -- (3.55,-0.95);
  \draw[arrow] (maj.south) -- (check.north west);
  \draw[arrow] (min.south) -- (check.north east);
  \node[align=center, text=qred] at (3.55,-3.15) {\textbf{\firstack{}: stale successful read}\\\textbf{\quorum{}: minority read unavailable}};
\end{tikzpicture}
\caption{Figure 1: the core stale-read mechanism. The violation is not replica divergence by
itself. The violation is a successful \firstack{} read of \texttt{v0} after a completed write to
\texttt{v1} must precede the read in every legal single-register order.}
\end{figure}

\section{Checker and Evidence Pipeline}

The checker consumes successful completed operations from a history and searches for a sequential
single-register order that preserves real-time precedence. If operation $a$ finishes before
operation $b$ starts, $a$ must appear before $b$ in any legal linearization. Writes update the
oracle value; reads must return the current oracle value at their chosen linearization point.

For linearizable histories, the checker returns a legal operation order and final register value.
For non-linearizable histories, it returns a witness when it can identify a stale read. It also
records diagnostic search state: checked operations, unavailable operations, real-time
predecessors, explored states, memoized dead ends, candidate operations, blocked candidates, and
reads rejected by the oracle value. This is why the workbench can expose the checker rather than
only display a final pass/fail verdict.

The product evidence pipeline has four layers:

\begin{enumerate}
  \item \textbf{Curated replay.} A split-brain stale-read fixture establishes a known minimal
  failure mode.
  \item \textbf{Adversarial search.} A seeded generator explores biased partition schedules,
  including overlapping client operations, and invokes the shrinker on failing scenarios.
  \item \textbf{Regression corpus.} Selected minimized failures and safe cases become fixtures
  with expected outcomes and provenance checks.
  \item \textbf{Tiny exhaustive exploration.} A small finite scenario model is enumerated to
  compare generated search with full coverage of a declared tiny space.
\end{enumerate}

\section{Counterexample Search, Shrinking, and Corpus Promotion}

The default adversarial search used for this report starts at seed 143 and explores 50 generated
schedules over 5 nodes, 8 operations, 3 clients, partition intensity 0.75, and concurrency
intensity 0.45. It found a \firstack{} violation at seed 143. The original generated scenario had
11 steps; the shrinker reduced it to 3 steps while preserving the checker failure:

\begin{quote}
\texttt{op2 read returned v0 after op1 write completed with v143-0-x1.}
\end{quote}

The important artifact behavior is not just finding a bug. It is preserving a small witness that
can be replayed, compared against \quorum{}, promoted into the corpus, and rechecked as part of
future product verification. The current corpus has four fixtures: a curated split-brain stale
read, a minimized generated search failure, a concurrent safe-overlap scenario, and a minimized
stale-read witness from the exhaustive explorer.

\section{Bounded Exhaustive Explorer}

The exhaustive explorer is intentionally tiny. Its purpose is not to prove a protocol generally
correct, but to replace vague search confidence with exact coverage for one finite model. The
current default model has 3 replicas, 2 clients, a single key, at most 3 operations, at most 2
topology changes, healed topology plus canonical 1/2 partitions, deterministic simulator timing,
and concurrent schedules. Message timing interleavings are not exhaustively enumerated.

The explorer checks every terminal history in that declared space under both protocols. It then
reports prefixes explored, terminal histories checked, unique scenarios, concurrent schedules,
partition shapes, operation patterns, violations, stale-read witnesses, and unavailable
operations. This gives a precise bounded claim: within this tiny model, \quorum{} produced zero
linearizability violations under the current assumptions. It remains explicitly not a proof for
arbitrary systems.

\section{Evaluation and Findings}

The reported results were captured from checkout \texttt{04ae315}. The current paper polish only
changes the report and PDF; it does not change product code or recompute expectations. Before this
revision, the product commands were rerun and the numbers below matched the CLI output.

\begin{table}[h]
\centering
\begingroup
\setlength{\tabcolsep}{3pt}
\scriptsize
\begin{tabular}{@{}L{0.20\linewidth}L{0.10\linewidth}L{0.09\linewidth}L{0.08\linewidth}L{0.10\linewidth}L{0.25\linewidth}@{}}
\toprule
\textbf{Evidence surface} & \textbf{Cases} & \textbf{FA viol.} & \textbf{Q viol.} & \textbf{Q unavail.} & \textbf{Finding} \\
\midrule
Regression corpus & 4 fixtures & 3/4 & 0/4 & 4 & Fixture expectations matched; no drift. \\
Adversarial search & 50 seeds & 50/50 & 0/50 & 155 & Generated schedules reliably recover stale reads. \\
Tiny exhaustive model & 1000 histories & 144 & 0 & 1064 & Full coverage of the declared tiny model agrees with the stale-read class. \\
Deterministic benchmark & 50 runs & 50/50 & 0/50 & 50 & Repeated 2/3 partition probes preserve the same safety/availability split. \\
\bottomrule
\end{tabular}
\endgroup
\caption{Summary of current findings. FA is \firstack{} and Q is \quorum{}. Counts are bounded to
the declared corpus, generated runs, benchmark runs, and tiny exhaustive model.}
\end{table}

\textbf{Finding 1: \firstack{} stale reads are robust in this model.} The stale-read class appears
in the curated fixture, in all 50 default adversarial generated schedules, in 3 of 4 regression
fixtures, in 144 terminal histories from the tiny exhaustive model, and in all 50 deterministic
benchmark runs.

\textbf{Finding 2: \quorum{} trades availability for safety in the checked corpora.} Across the
reported surfaces, \quorum{} produced zero linearizability violations, but it returned
unavailable for minority-side operations: 4 operations in the corpus, 155 in the generated search
comparison, 1064 in the tiny exhaustive model, and 50 in the benchmark.

\textbf{Finding 3: generated search and tiny exhaustive exploration answer different questions.}
Adversarial search samples larger biased schedules and quickly finds compact failures. Exhaustive
exploration covers every scenario in a much smaller declared model and provides exact coverage
accounting. Agreement between the two strengthens confidence in the modeled failure class, but
neither result is a general protocol proof.

\section{Workbench Surface}

The React/Vite workbench is subordinate to the engine. It exists to expose the technical object:
the generated or loaded scenario, the protocol comparison, the partition replay, the operation
timeline, the event trace, the checker verdict, minimized counterexamples, and oracle
diagnostics. The most important UI addition is the oracle trace panel: it shows real-time
predecessor constraints and captured checker states so that a user can see why a read was
accepted, blocked, or rejected.

The CLI remains the stronger research surface because it produces stable evidence: product
reports, corpus validation, generated search summaries, exhaustive coverage accounting, benchmark
results, and reproduction commands for concrete witnesses.

\section{Limitations}

\qs{} is intentionally narrow. The limitations are part of the artifact's credibility:

\begin{itemize}
  \item The object is a single-key register, not a multi-key database or transaction system.
  \item The network model is reachability under partitions, not a real network stack.
  \item Search is bounded and generated; exhaustive exploration covers only the declared tiny
  finite model.
  \item Simulator timing is deterministic and does not enumerate all message-delay schedules.
  \item The included protocols are simple \firstack{} and majority \quorum{} register protocols,
  not Raft, Paxos, leader election, or production replication.
  \item The model excludes durability, crash recovery, retries, read repair, Byzantine faults,
  and deployment behavior.
  \item The absence of \quorum{} violations is bounded evidence for this model and these corpora,
  not a formal proof.
\end{itemize}

\section{Research Directions}

The natural next step is not a broader UI. It is a stronger model. Useful directions include a
richer scheduler and message-delay model, larger but still explicit exhaustive bounds, multi-key
histories, predicate reads, transaction-isolation variants, and additional protocol models once
the register model remains stable under those stronger checks. Each extension should preserve the
artifact discipline used here: explicit assumptions, deterministic fixtures, corpus expectations,
shrunk witnesses, bounded claims, and reproducible result capture.

\appendix
\section{Result Capture Notes}

The verification process for this report used the local product commands that exercise the test
suite, product report, deterministic benchmark, adversarial search, exhaustive explorer, and
corpus runner. The README remains the command reference. In this report, reproducibility is
represented by the evidence trail: the checkout hash \texttt{04ae315}, fixed seeds, fixture
hashes in the corpus output, minimized counterexample identifiers, and the bounded claim emitted
by the exhaustive explorer.

\end{document}
